Resource Buying Games

### 2017-01-11 slides.tex
added lemma and theorem for non-matroid space

@@

\begin{frame}
  \frametitle{Soundness}
  \begin{itemize}
    \item Termination:
      \begin{itemize}
        \item inductively applying contraction operations lead to basis
        \item deletion cannot \enquote{loose} bases for other players since
          chosen cut is inclusion-wise minimal
      \end{itemize}
    \item Correctness
      \begin{itemize}
        \item Lemma: cost of the chosen element decreases monotonically in each
          iteration
        \item assume there is a beneficial alternative basis
          $\hat{B} = (B_{i}\setminus\set{g})\cup{f}$
          \begin{itemize}
            \item consider iteration $k$ in which $g$ is added to basis of $i$:
            \item if $f$ is in the corresonponding cut its cost never changes
              since it is deleted from every matroid $rightarrow$ $f$ is not in
              the cut
            \item but $f$ and $g$ \enquote{open} same dimensionality, thus $f$
              is not left in contracted matroid of $i$
            \item $f$ was deleted before (in this iteration the cost of the
              chosen element was lower than for $f$ in contradiction to lemma)
          \end{itemize}
      \end{itemize}
  \end{itemize}
\end{frame}

\section{Weighted Matroid Games}
\begin{frame}
  % \begin{itemize}
  %   \item (2 slides)
  %   \item Introduce socially optimal solution
  %   \item Show (or at least motivate) why this is a PNE
  %   \item Mention reduction to show that finding an optimal solution is
  %     NP-hard
  % \end{itemize}
\end{frame}

\section{Non-Matroid Strategy Spaces}
\subsection{Lack of PNE}
\begin{frame}
  % \begin{itemize}
  %   \item (1 slide)
  %   \item Present that for Non-Matroid Games we can find strategy spaces
  %     which do not allow PNE
  %   \item Illustrate the construction by a picture
  % \end{itemize}
\end{frame}

\subsection{Non-Decreasing Marginal Cost Functions \& Price of Anarchy}
\begin{frame}
  % \begin{itemize}
  %   \item (3 slides)
  %   \item Present that games with Non-Decreasing Marginal Cost Functions allow
  %     PNE (explain how the marginal cost pricing is obtained)
  %   \item Define Price of Anarchy
  %   \item Mention that non-increasing and non-decreasing games with weighted
  %     players is unbounded
  % \end{itemize}
\end{frame}

\section{Conclusion}
\begin{frame}
  % \begin{itemize}
  %   \item (1 slide)
  %   \item Sum up results
  %   \item Embed results into topic
  % \end{itemize}
\end{frame}

>>>>

\DeclareMathOperator{\ex}{ex}

@@

\begin{frame}
  \frametitle{Soundness}
  \begin{itemize}
    \item Termination:
      \begin{itemize}
        \item inductively applying contraction operations lead to basis
        \item deletion cannot \enquote{loose} bases for other players since
          chosen cut is inclusion-wise minimal
      \end{itemize}
    \item Correctness
      \begin{itemize}
        \item Lemma: cost of the chosen element decreases monotonically in each
          iteration
        \item assume there is a beneficial alternative basis
          $\hat{B} = (B_{i}\setminus\set{g})\cup{f}$
          \begin{itemize}
            \item consider iteration $k$ in which $g$ is added to basis of $i$:
            \item if $f$ is in the corresonponding cut its cost never changes
              since it is deleted from every matroid $\rightarrow$ $f$ is not
              in the cut
            \item but $f$ and $g$ \enquote{open} same dimensionality, thus $f$
              is not left in contracted matroid of $i$
            \item $f$ was deleted before (in this iteration the cost of the
              chosen element was lower than for $f$ in contradiction to lemma)
          \end{itemize}
      \end{itemize}
  \end{itemize}
\end{frame}

\section{Weighted Matroid Games}
\begin{frame}
  \frametitle{Recall: Congestion Model}
  \begin{equation*}
    \mathcal{M} = \tupel{N, E, \mathcal{B}, \left(d_{i}\right)_{i\in N},
    \left(c_{e}\right)_{e\in E}}
  \end{equation*}
  \vspace{-1cm}
  \begin{itemize}
    \item before: $d_{i} = 1$ for all $i$
    \item now: $d_{i}\in\mathbb{N}_{>0}$ 
    \item auxiliary definitions:
      \begin{itemize}
        \item $N_{e}\tupel{B} = \set{i\in N\middle|e\in B_{i}}$
        \item for a fixed $B$: $\ex_{i}\tupel{e} =
          \set{f\in E\setminus\set{e}\middle|
          (B_{i}\setminus{e})\cup\set{f}\in\mathcal{B}_{i}}$: set of resources
          that can replace $e$ for player $i$
        \item $F = \bigcup_{i\in N}\bigcap_{B\in \mathcal{B}_{i}} B$:
          set of obligatory elements for at least one player
        \item $\Delta_{i}\tupel{B;e\rightarrow f} =
          c_{f}\tupel{\ell_{f}\tupel{B_{i}\cup\set{f}\setminus\set{e}}} -
          c_{f}\tupel{\ell_{f}\tupel{B}}$: marginal cost to buy resource $f$ if
          $i$ switches
        \item $\Delta_{i}^{e} = \min_{f\in\ex_{i}(e)}\set{\Delta_{i}(B;e\rightarrow f)}$:
          minimal marginal cost for exchanging $e$
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Characterisation of Pure Nash Equilibria}
  \begin{theorem}
    For a weighted matroid resource buying game
    $\tupel{N, E, \mathcal{B}, d, c}$ and $B\in\mathcal{B}$ the following
    holds: there is a payment vector $p$ such that $\tupel{B,p}$ is a
    PNE if and only if
    \begin{equation*}
      c_{e}(B) \leq \sum_{i\in N_{e}(B)}\Delta_{i}^{e}(B)
      \textit{ for all }e\in E\setminus F
    \end{equation*}
  \end{theorem}
  \begin{proof}
    if: $p_{i}^{e} =
      \begin{cases}
        \frac{\Delta_{i}^{e}(B)}{\sum_{j\in N_{e}(B)}\Delta_{j}^{e}(B)}\cdot
        c_{e}(B)&\text{ if } \sum_{j\in N_{e}(B)}\Delta_{j}^{e}(B) > 0\\
        0 &\text{ otherwise}
      \end{cases}$\\
      for all $e\in E\setminus F$ and $i\in N_{e}(B)$\\
    only if: for all $i$ and $e$: $p_{i}^{e} \leq \Delta_{i}^{e}(B)$
      and $c_{e}(B) = \sum_{i\in N_{e}(B)}p_{i}^{e}$
  \end{proof}
\end{frame}

\begin{frame}
  \frametitle{Existence of Pure Nash Equilibria}
  \begin{theorem}
    Every weighted matroid resource buying game with marginally non-increasing
    cost functions possesses a pure Nash equilibria.
  \end{theorem}
  \begin{itemize}
    \item socially optimal configuration profile: $B\in\mathcal{B}$ such that
      $\sum_{e\in E}c_{e}(\ell_{e}(B))$ is minimal
    \item for socially optimal configuration profile $B$ holds:
      \begin{equation*}
        c_{e}(B)\leq \sum_{i\in N_{e}(B)}\Delta_{i}^{e}(B)
        \text{ for all }e\in E\setminus F
      \end{equation*}
    \item $B$ is a pure Nash equilibrium
    \item (but finding socially optimal configuration profiles is NP-hard)
  \end{itemize}
\end{frame}

\section{Non-Matroid Strategy Spaces}
\begin{frame}
\frametitle{Consequences of using Non-Matroid Strategy Spaces}
  \begin{lemma}
    If $\mathcal{S}\subseteq 2^{E}$ is a non-matroid anti-chain, then there
    exist $X,Y\in\mathcal{S}$ and $\set{a,b,c}\subseteq X\cup Y$ such that each
    set $Z\in\mathcal{S}$ with $Z\subseteq(X\cup Y)\setminus\set{a}$ contains
    $b$ and $c$. (Especially, there is $S\in\mathcal{S}$ with
    $S\subseteq X\cup Y$ and $a\in S$ but $b,c\notin S$.)
  \end{lemma}
  \begin{theorem}
    For every non-matroid anti-chain $\mathcal{S}$ on a set of resource $E$,
    there exists a weighted two-player resource buying game
    $G = \tupel{\tilde{E}, \tupel{\mathcal{S}_{1} \times
    \mathcal{S}_{2}}\times \mathcal{P}, \pi}$ having marginally non-increasing
    cost functions, whose strategy spaces $\mathcal{S}_{1}$ and
    $\mathcal{S}_{2}$ are both isomophic to $\mathcal{S}$, so that $G$ does not
    possess a pure Nash equilibrium.
  \end{theorem}
\end{frame}

\begin{frame}
  \frametitle{Construction of games without PNE}

\end{frame}

\section{Conclusion}
\begin{frame}
\end{frame}
